%!TEX root = ../thesis.tex

%I would like to thank Jonathan Polimeni for cleaning up all these old
%LaTeX style files and templates so that I didn't have to use Word to
%write up this document. Also, I would like to thank all the CV/CNS lab
%graduates for their contributions and tweaks to this organizational
%scheme over the years (after many dissapointing interactions with
%Martha Wellman).
%
%This brings me to thank Martha Wellman of CAS and Brendon McDermot of
%Mugar library who together uphold the stylistic and aesthetic
%conventions that have been implemented in this LaTeX manuscript. Also
%Sister Mary Virginia, the Thesis/dissertation coordinator of Mugar
%Library, deserves some thanks as well.
%
%Finally, I would like to thank Stephen Gildea, for the MIT sytle file
%off which this current version is based, and Paolo Gaudiano for
%porting the MIT style to one compatible for BU.

% [[[ LORENA ]]]
I would like to express wholehearted gratitude to my academic advisor over the last 6 years, Prof. Barba for her support and guidance throughout my studies in two universities, 3 degrees and 2 continents. In particular I would like to thank her for the belief she showed in me by bringing me from Bristol to Boston University in order to continue my studies. I deeply appreciate the freedom I have been given in my research, along with the invaluable advice and support that has been key to my successes.

% [[[ COMMITTEE ]]]
I would also like to thank the members of my committee, Professors Paul Barbone, Sheryl Grace and Emily Ryan for their time and insights during these final months of my studies.

% [[[ GROUP ]]]
With the contributions, ideas and friendship of Christopher Cooper, Anush Krishnan, Brent Parham and Olivier Mesnard in Prof. Barba's research group, the time I have spent here has flown by, and they have always been willing to help in any way necessary, be it debugging code, reading papers or spending time relaxing away from the office. Special thanks must also go to Rio Yokota, who was always available for advice on coding, {\fmm} or any other subject.

% [[[ OTHERS ]]]
I also take this opportunity to thank those who have helped and funded me through my time at Boston University -- Dan Kamalic for his patience in helping me with the use of his compute clusters. My continued funding has been greatly appreciated, and I would like to thank Boston University, the Massachusetts Green High-Performance Computing Center (MGHPCC), National Science Foundation (NSF) and the Office of Naval Research (ONR) for their contributions. 

% [[[ PARENTS ]]]
I would like to thank my fianc\'ee, Elizabeth Matos for the love and support she has shown me, both pulling me through the last parts of my studies, and giving me an incredible future to look forward to. Finally, I wish to show my appreciation for my parents, Dennis and Wendy Layton for their unending and unconditional encouragement and belief, for always being there for me, and for not begrudging me this time far away from them in Boston. 
